---
title: "HW3: Tidy Data"
author: "Jenna Loesberg"
date: "10/6/2019"
output: html_document
---

[Instructions are found here](https://stat545.stat.ubc.ca/evaluation/hw04/hw04/)

```{r , warning=FALSE,echo=FALSE}
suppressPackageStartupMessages(library(tidyverse)) 
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(DT))
```

### Exercise 1:
#### Univariate Option 1:
#### 1.1. Make a tibble with one row per year, and columns for life expectancy for two or more countries.

What it looks like before:
```{r}
gapminder %>% 
  datatable()
```

Using pivot_wider to get table where each year is a new entry, and include life expectancy for 2 countries:
```{r}
gapminder2 <- 
  gapminder %>% 
  filter(country == "Australia"| country ==  "Canada") %>% 
  select(country, year, lifeExp) %>% 
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = lifeExp)

datatable(gapminder2)

```

#### 1.2. Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.

```{r}
gapminder2 %>% 
  ggplot() +
  geom_point(aes(Australia, Canada))+
  xlab("Life Expectancy in Australia") +
  ylab("Life Expectancy in Canada")
```


#### 1.3. Re-lengthen the data.

```{r}
gapminder2 %>% 
   pivot_longer(cols = c(Australia, Canada),
               names_to = "country",
               values_to = "lifeExp") %>% 
  arrange(country) %>% 
  datatable()
```


### Exercise 2
#### Multivariate Option 1
#### 2.1. Make a tibble with one row per year, and columns for life expectancy and GDP per capita (or two other numeric variables) for two or more countries.

```{r}
gapminder3 <- 
  gapminder %>% 
  filter(country == "Australia"| country ==  "Canada") %>% 
  select(country, year, lifeExp, gdpPercap) %>% 
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = c(lifeExp, gdpPercap))
datatable(gapminder3)
```

#### 2.2. Re-lengthen the data.
```{r}
gapminder3 %>% 
  pivot_longer(cols = -year,
               names_to = c(".value", "country"),
               names_sep = "_") %>%
  arrange(country) %>% 
  datatable()
```
###  Exercise 3

reading in data
```{r, message = FALSE, warning = FALSE}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

#### 3.1
For each guest in the guestlist (guest tibble), add a column for email address, which can be found in the email tibble.

Need to rearrange the email file so it matches the guest file:
```{r}
email2 <- 
  email %>%
  separate_rows(guest, sep = ", ") %>%
  rename(name = guest)
 
guest %>% 
  left_join(email2, by = c("name")) %>% 
  datatable()
guest
```
Great, now each person/ group has an email and its all on one file

#### 3.2

Who do we have emails for, yet are not on the guestlist?

```{r}

email2 %>%
  anti_join(guest, by = c("name")) %>% 
  datatable()


```

#### 3.3
Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist). (adding the three extra people to the guest list)
```{r}
full_list <- guest %>%
  full_join(email2, by = c("name")) %>% 
  datatable()
full_list
```
And there they are!