---
title: "HW02"
author: "Jenna Loesberg"
date: "9/23/2019"
output: github_document
---
Instructions are provided [here](https://stat545.stat.ubc.ca/evaluation/hw02/hw02/) 

Loading packages:
```{r load packages, warning = FALSE, message = FALSE}
library(gapminder)
library(tidyverse)
```
# Exercise 1
## 1.1
Using filter() to get data for 3 countries in 1970's
```{r}
gapminder %>%
  filter(country == "United States" | country == "Norway" | country == "New Zealand") %>%
  filter(year >= 1970 & year <= 1979)
```

## 1.2
using pipe operator %>%  to select country and gdpPercap
```{r}
gapminder %>%
  filter(country == "United States" | country == "Norway" | country == "New Zealand") %>%
  filter(year >= 1970 & year <= 1979) %>% 
  select(country, gdpPercap)
```

## 1.3
Filtering for drops in life expectancy (LE) and adding new variable for change in LE (I'm filtering for drops in LE, so these will be negative)
```{r}
(gapminder2 <- 
   gapminder %>% 
   arrange(year) %>% 
   group_by(country) %>% 
   mutate(LE_change = lifeExp - lag(lifeExp)) %>% 
   filter(LE_change < 0))
```
## 1.4
Showing max GDP per capita experienced by each country:
```{r}
gapminder %>%
  group_by(country) %>% 
  filter(gdpPercap == max(gdpPercap))
```

## 1.5
Making scatterplot of log transformed GDP and life expectancy in Canada
```{r}
gapminder %>%
  filter(country == "Canada") %>% 
  ggplot(aes(gdpPercap, lifeExp))+
  geom_point()+
  scale_x_log10()+
  xlab("log(GDP per capita)")+
  ylab("Life Expectancy (years)")
```

# Exercise 2
### Exploring individual variables with dyplr

Categorical: __continent__ 
Quantitative: __lifeExp__

## 2.1
What are possible values (or range, whichever is appropriate) of each variable?

```{r}
summary(gapminder)
```
#### Life Expectancy: 
Ranges from 23.60 to 82.60; Mean is 59.47... see summary table

Frequency of variables are skewed to the right...
```{r}
hist(gapminder$lifeExp)
```

#### Continent:
5 categories, but that summary table is misleading because it includes all of the years that data was collected. So, I'm isolating it to one year:
```{r}
(gapminder3 <- 
  gapminder %>% 
  filter(year == "2002"))
```

Now the real summary...

```{r}
summary(gapminder3)
```
Here we see the number of countries from each continent.

# Exercise 3
### Making graphs

## 3.1
A scatterplot of 2 quantitative variables:

comparing population and GDP in China...does gdp increase as population increases? This spans from 1952 to 2007
```{r}
gapminder %>% 
  filter(country == "China") %>%
  ggplot(aes(pop, gdpPercap))+
    geom_point()+
    xlab("Population")+
    ylab("GDP per capita")
```

## 3.2
A boxplot of life expectancy per continent in 2002:

```{r}
gapminder %>% 
  filter(year == "2002") %>%
  ggplot(aes(continent, lifeExp))+
  geom_boxplot()+
  xlab("Continent")+
  ylab("Life expectancy (years)")
```

